### 2022-04-22 greta/hopf_mcmc.Rmd
mcmc ...

```{r}
step <- function(H, P, eta,  t=0, 
                 p = list(r = 0.75, c = 0.25, Ko = 15, delta=0.1)) {
  K <- p$Ko + p$delta*t
  N <- numeric(2)
  N[1] <- H * exp(p$r * (1 - H/K) - p$c * P + eta[1])
  N[2] <- H * exp(p$r * (1 - H/K) ) * (1 - exp(-p$c * P+ eta[2]) )
  N[1] <- np.clip(N[1], 0, 100)
  N[2] <- np.clip(N[2], 0, 100)
  N
}

simulate <- function(t_max = 100, 
                     p = list(H_init = 9, P_init = 1,
                              r = 0.75, c = 0.1,  Ko = 15, delta = 0.1,
                              sigma_H = 2e-2, sigma_P = 2e-2)
                     ) {
  eta <- array(c(rnorm(t_max, 0, p$sigma_H), 
                 rnorm(t_max, 0, p$sigma_P)),
               dim=c(t_max,2))
  N   <- array(NA, c(t_max,2))

@@

```{r}
t_max <- train_t_max + test_t_max
 p = list(H_init = 9, P_init = 1,
          r = 0.75, c = 0.1,  Ko = 14, delta = 0.08,
          sigma_H = 2e-2, sigma_P = 2e-2)
sim <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=t_max, p = p), .id = "i")
train <- sim |> filter(t <= train_t_max)
test <- sim |> filter(t > train_t_max)
# test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

>>>>

```{r}
step <- function(H, P, eta,  t=0, p) {
  K <- p$Ko + p$delta*(t+p$t_init)
  N <- numeric(2)
  N[1] <- H * exp(p$r * (1 - H/K) - p$c * P + eta[1])
  N[2] <- H * exp(p$r * (1 - H/K) ) * (1 - exp(-p$c * P+ eta[2]) )
  N[1] <- np.clip(N[1], 0, 100)
  N[2] <- np.clip(N[2], 0, 100)
  N
}

simulate <- function(t_max = 100, 
                     p = list(H_init = 9, P_init = 1,
                              r = 0.75, c = 0.1,  Ko = 15, delta = 0.1,
                              sigma_H = 2e-2, sigma_P = 2e-2, t_init = 0)
                     ) {
  eta <- array(c(rnorm(t_max, 0, p$sigma_H), 
                 rnorm(t_max, 0, p$sigma_P)),
               dim=c(t_max,2))
  N   <- array(NA, c(t_max,2))

@@

```{r}
t_max <- train_t_max + test_t_max
 p = list(H_init = 9, P_init = 1,
          r = 0.75, c = 0.1,  Ko = 14, delta = 0.08,
          sigma_H = 2e-2, sigma_P = 2e-2, t_init = 0)
sim <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=t_max, p = p), .id = "i")
train <- sim |> filter(t <= train_t_max)
test <- sim |> filter(t > train_t_max)
# test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

### 2022-04-22 greta/sn_mcmc.Rmd
mcmc ...

set.seed(4242)
train_reps <- 1
train_t_max <- 75
test_t_max <- 75
test_reps <- 100

@@

        h <- np.clip(pars$h_init + pars$alpha * t, 0, 0.27)
        N <- N + pars$r * N * (1 - N / pars$K) - 
          h * (N**2 / (pars$s**2 + N**2)) + eta
        N <- np.clip(N, 0, 1)
        N
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 250L,
                     pars = list(
                        r = 1,
                        K = 1,
                        s = 0.1,
                        h_init = .15,
                        alpha = 0.0015,
                        mu = 0,
                        sigma = 0.00
                      )
                    ) {
  eta <- rnorm(t_max, pars$mu, pars$sigma) # mu = 0, no drift
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], t, pars)
  }
  tibble::tibble(t = 1:t_max, N = N)
}
```

```{r}
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02
          )
t_max <- train_t_max + test_t_max
sims <- purrr::map_dfr(1:train_reps, 
                       \(i) simulate(t_max=t_max, pars = pars), 
                       .id = "i")

@@

library(greta)
r <- uniform(0, 10)
K <- uniform(0, 10)
s <- uniform(0,10)
h_init <- uniform(0,10)
alpha <- uniform(0, 10)
sigma <- uniform(0, 10)
h <- h_init + alpha * t
mean <-  x_t + r*x_t*(1 - x_t / K) - h*(x_t^2 / (s^2 + x_t^2))
distribution(x_t1) <- normal(mean, sigma)
m <- model(r, K, s, h_init, alpha, sigma)
```

@@

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 60000, warmup = 50000, chains = 4, verbose = FALSE)
})
```

@@

```{r}
## draw test_reps number of samples

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps)
posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max, p = q) ,.id = "i")

bind_rows(
  mutate(test, model="true"), 
  mutate(posterior_sims, model="predicted")
)|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.1)

@@

```{r}
true <- tibble(a = 0.75, b = 0.25, sigma = 1e-2) |> gather(variable, value)
bind_rows(map(draws, as_tibble)) |>
  gather(variable, value) |> ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

>>>>

set.seed(4242)
train_reps <- 100
train_t_max <- 75
test_t_max <- 75
test_reps <- 100

@@

        h <- np.clip(pars$h_init + pars$alpha * (pars$t_init + t), 0, 0.27)
        N <- N + pars$r * N * (1 - N / pars$K) - 
          h * (N**2 / (pars$s**2 + N**2)) + eta
        N <- np.clip(N, 0, 100)
        N
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 250L,
                     pars = list(
                        r = 1,
                        K = 1,
                        s = 0.1,
                        h_init = .15,
                        alpha = 0.0015,
                        mu = 0,
                        sigma = 0.00,
                        t_init = 0
                      )
                    ) {
  eta <- rnorm(t_max, pars$mu, pars$sigma) # mu = 0, no drift
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], t, pars)
  }
  tibble::tibble(t = pars$t_init + 1:t_max, N = N)
}
```

```{r}
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02,
            t_init = 0
          )
t_max <- train_t_max + test_t_max
sims <- purrr::map_dfr(1:train_reps, 
                       \(i) simulate(t_max=t_max, pars = pars), 
                       .id = "i")

@@

library(greta)
r <- uniform(0, 10)
K <- uniform(0, 10)
s <- uniform(0, 10)
h_init <- uniform(0, 10)
alpha <- uniform(0, 10)
sigma <- uniform(0, 10)
h <- h_init + alpha * t
mean <-  x_t + r*x_t*(1 - x_t / K) - h*(x_t^2 / (s^2 + x_t^2))
distribution(x_t1) <- normal(mean, sigma)
m <- model(r, K, s, h_init, alpha, sigma)
```

@@

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 100000, warmup = 80000, chains = 4, verbose = FALSE)
})
```

@@

```{r}
## draw test_reps number of samples
left_off <- train |> 
  filter(t == train_t_max) |> 
  select(N) |> 
  rename(N_init = N) |>
  slice_sample(n = test_reps, replace=TRUE)

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) |> 
  sample_n(test_reps) |>
  mutate(t_init = train_t_max) |>
  bind_cols(left_off)

posterior_sims <- posterior_samples |>
  mutate(mu=0) |>
  purrr::transpose() |>
  map_dfr(function(q) simulate(t_max = test_t_max, pars = q, N_init = q$N_init) ,.id = "i")

bind_rows(
  mutate(train, model="historical"),
  mutate(test, model="true"), 
  mutate(posterior_sims, model="predicted")
)|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.2)

@@

```{r}
true <- as_tibble(pars) |> gather(variable, value)
bind_rows(map(draws, as_tibble)) |>
  gather(variable, value) |> ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```
